A modern, event-based architecture for distributed evolutionary algorithms

### 2018-04-04 paper.tex
last changes

Cloud native applications add a layer of abstraction to the
underlying distributed computing system, defining a high-level,
self-scaling and self-managed architecture of different microservices
linked by a messaging bus. Creating new algorithms that tap these
architectural patterns and at the same time employ distributed
resources efficiently is a challenge we will be taking up in this
paper. We introduce KafkEO, a cloud native evolutionary algorithms
framework that is prepared to work with different implementations of evolutionary
algorithms and other population-based metaheuristics by using
micro-populations and stateless services as the main building
blocks; KafkEO is an attempt to map the traditional evolutionary
algorithm to this new cloud-native format. As far as we know, this is the first
architecture of this kind that has been published and tested, and is free 
software and vendor-independent, based on OpenWhisk and Kafka. 
This paper presents a proof of concept, examines its cost, and tests 
the impact on the algorithm of the design around
cloud native and asynchronous system by comparing it on
the well known BBOB benchmarks with other pool-based architectures,
with which it has remarkable functional resemblance. KafkEO results are 
quite competitive with similar architectures.

@@

\section{Introduction}
Cloud computing is increasingly becoming the dominant way of running
the server side of most enterprise applications nowadays, the same as the browser is
the standard platform for the client side. Besides the convenience of the pay-as-you-go
model, it also offers a way of describing the infrastructure as part
of the code, so that it is much easier to reproduce results and has 
been a boon for scientific computing.
%The availability of resources only limited by your budget, and in many
%cases for free, as well as the inherent reproducibility of cloud
%deployments, has been a boon for scientific computing. 
However,
programming the cloud means that monolithic applications, that is,
applications built on a single stack of services that communicate by
layers, are no longer an efficient architectural design for scientific workflows. Cloud
architectures favor asynchronous communication over heterogeneous %favor: american
resources, and shifting from mostly sequential and monolithic 
to an asynchronously parallel architecture will also imply important
reformulation of the algorithms in order to take full advantage of these technologies.
  Cloud native applications add a layer of abstraction to the
  underlying distributed computing system, seamlessly integrating
  different elements in a single data flow, allowing the user to just focus on code and service connections. Services are {\em native} points
  in this new architecture, departing from a monolithic or even
  distributed paradigm to become a loosely collection of
services, in fact {\em microservices} \cite{microservices}, which in
 many cases are stateless, reacting to some event and
  {\em living} only while they are doing some kind of processing.
%   and not keeping state from one invocation to the next.
% redundant - Mario  
  Reactive systems not only allow massive scaling and independent deployment they are also more economical than other monolithic options.  Platform as a service (PaaS) or even Container as a Service (CaaS) approaches need to be running all
  the time in order to maintain their state, so they are paid for their size and
  time they remain active.
  At any rate, while one of the main selling points of Functions as a Service (FaaS) is their ultra-fast activation time, from our point of view their most interesting feature is the fact that they provide stateless processing. An important caveat
  of stateless processing is that algorithms must be adapted to this fact and
  turned, at least in part, into a series of stateless steps thaworking on a data stream.
    It is also taken to an atomic extreme
    with the so-called serverless architectures \cite{Varghese2018849},
    which allow vendors and users to deploy code as single, stateless
    functions, that get activated via {\em rules}, {\em triggers} or
    explicitly, reacting to events consumed from
    a message queue.
    %or bus that
    % generally reacting to events from a message queue or bus that
    % what is generated? rules are triggered and triggers fire in
    % response to events
    % Please change it to whatever you think it's the best - JJ
    %is the backbone of the application. 
    The first commercial
    implementation of this kind of architecture was released by Amazon
    with its Lambda product, to be closely followed by releases by Azure
    (Microsoft) and Google and OpenWhisk, an open source implementation
    released by IBM \cite{Baldini2016287}.

@@

  The experiments using the above described framework were performed
  during the month of January using a paid IBM BlueMix subscription. The
  totality of experiments costed about \$12. Most of the cost is due to
  the MessageHub {\em partitions}, that is, the hosted messaging service
  itself. The amount paid for the messages in the BlueMix platform
  is less than one dollar in total; messages are paid by the hundreds of
  thousands delivered, and are actually not the most expensive part of the
  implementation of the algorithm.
  Partitions are
  essential for a high throughput; a messaging queue will be able to
  process as many messages as the partitions are able to get through in
  parallel; this means that cost will scale with the number of messages
  in a complex way, not simply linearly, and design decisions will have
  to be taken. The baseline is that the best option is to maximize the
  number of messages that can be borne by a particular partition, and
  try to minimize the number of partitions to avoid scaling costs.

  \begin{figure*}[h!tb]
  \begin{tabular}{l@{\hspace*{-0.025\textwidth}}l@{\hspace*{-0.025\textwidth}}l@{\hspace*{-0.025\textwidth}}l}
  \includegraphics[width=0.266\textwidth]{ppfigdim_f001}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f006}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f002}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f007}\\
  \includegraphics[width=0.266\textwidth]{ppfigdim_f003}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f008}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f004}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f009}\\
  \includegraphics[width=0.266\textwidth]{ppfigdim_f005}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f010}\\
  \end{tabular}
  \vspace{-3ex}
   \caption{\label{fig:aRTgraphs}
   \bbobppfigdimlegend{$f_1$ and $f_{24}$}
  }
  \end{figure*}
  %
  The results of the comparison are shown in Figure \ref{fig:aRTgraphs},
  which follow the classical BBOB 2009 format, which includes the amount
  of effort devoted to finding a certain fitness level and time needed
  to do it. The EvoSpace and KafkEO results are shown side by side, only
  for the sake of comparison, since we are only interested for the time
  being in the baseline performance of the proof of concept.

>>>>

Cloud-native applications add a layer of abstraction to the
underlying distributed computing system, defining a high-level,
self-scaling and self-managed architecture of different microservices
linked by a messaging bus. Creating new algorithms that tap these
architectural patterns and at the same time employ distributed
resources efficiently is a challenge we will be taking up in this
paper. We introduce KafkEO, a cloud-native evolutionary algorithms
framework that is prepared to work with different implementations of
evolutionary algorithms and other population-based metaheuristics by using
micro-populations and stateless services as the main building
blocks; KafkEO is an attempt to map the traditional evolutionary
algorithm to this new cloud-native format. As far as we know, this is the first
architecture of this kind that has been published and tested, and is free 
software and vendor-independent, based on OpenWhisk and Kafka. 
This paper presents a proof of concept, examines its cost, and tests 
the impact on the algorithm of the design around
cloud-native and asynchronous system by comparing it on
the well known BBOB benchmarks with other pool-based architectures,
with which it has a remarkable functional resemblance. KafkEO results are 
quite competitive with similar architectures.

@@

\section{Introduction}
Cloud computing is increasingly becoming the dominant way of running
the server side of most enterprise applications nowadays, the same as the browser is
the standard platform for the client side. Besides the convenience of the pay-as-you-go
model, it also offers a way of describing the infrastructure as part
of the code, so that it is much easier to reproduce results and has 
been a boon for scientific computing.
%The availability of resources only limited by your budget, and in many
%cases for free, as well as the inherent reproducibility of cloud
%deployments, has been a boon for scientific computing. 
However,
programming the cloud means that monolithic applications, that is,
applications built on a single stack of services that communicate by
layers, are no longer an efficient architectural design for scientific workflows. Cloud
architectures favor asynchronous communication over heterogeneous %favor: american
resources, and shifting from mostly sequential and monolithic 
to an asynchronously parallel architecture will also imply important
reformulation of the algorithms in order to take full advantage of these technologies.
  Cloud-native applications add a layer of abstraction to the
  underlying distributed computing system, seamlessly integrating
  different elements in a single data flow, allowing the user to just focus on code and service connections. Services are {\em native} points
  in this new architecture, departing from a monolithic or even
  distributed paradigm to become a loosely collection of
services, in fact {\em microservices} \cite{microservices}, which in
 many cases are stateless, reacting to some event and
  {\em living} only while they are doing some kind of processing.
%   and not keeping state from one invocation to the next.
% redundant - Mario  
  Reactive systems not only allow massive scaling and independent deployment they are also more economical than other monolithic options.  Platform as a service (PaaS) or even Container as a Service (CaaS) approaches need to be running all
  the time in order to maintain their state, so they are paid for their size and
  time they remain active.
  At any rate, while one of the main selling points of Functions as a Service (FaaS) is their ultra-fast activation time, from our point of view their most interesting feature is the fact that they provide stateless processing. An important caveat
  of stateless processing is that algorithms must be adapted to this fact and
  turned, at least in part, into a series of stateless steps thaworking on a data stream.
    It is also taken to an atomic extreme
    with the so-called serverless architectures \cite{Varghese2018849},
    which allow vendors and users to deploy code as single, stateless
    functions, that get activated via {\em rules}, {\em triggers} or
    explicitly, reacting to events consumed from
    a message queue.
    %or bus that
    % generally reacting to events from a message queue or bus that
    % what is generated? rules are triggered and triggers fire in
    % response to events
    % Please change it to whatever you think it's the best - JJ
    %is the backbone of the application. 
    The first commercial
    implementation of this kind of architecture was released by Amazon
    with its Lambda product, to be closely followed by releases by Azure
    (Microsoft) and Google and OpenWhisk, an open source implementation
    released by IBM \cite{Baldini2016287}.

@@

  The experiments were performed
  during the month of January using a paid IBM BlueMix subscription. The
  totality of experiments costed about \$12. Most of the cost is due to
  the MessageHub {\em partitions}, that is, the hosted messaging service
  itself. The amount paid for the messages in the BlueMix platform
  is less than one dollar in total; messages are paid by the hundreds of
  thousands delivered, and are actually not the most expensive part of the
  implementation of the algorithm.
  Partitions are
  essential for a high throughput; a messaging queue will be able to
  process as many messages as the partitions are able to get through in
  parallel; this means that cost will scale with the number of messages
  in a complex way, not simply linearly, and design decisions will have
  to be taken. The baseline is that the best option is to maximize the
  number of messages that can be borne by a particular partition, and
  try to minimize the number of partitions to avoid scaling costs.

  \begin{figure*}[h!tb]
  \begin{tabular}{l@{\hspace*{-0.025\textwidth}}l@{\hspace*{-0.025\textwidth}}l@{\hspace*{-0.025\textwidth}}l}
  \includegraphics[width=0.266\textwidth]{ppfigdim_f001}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f006}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f002}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f007}\\
  \includegraphics[width=0.266\textwidth]{ppfigdim_f003}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f008}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f004}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f009}\\
  \includegraphics[width=0.266\textwidth]{ppfigdim_f005}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f010}\\
  \end{tabular}
  \vspace{-3ex}
   \caption{\label{fig:aRTgraphs}
   \bbobppfigdimlegend{$f_1$ and $f_{24}$}
  }
  \end{figure*}
  %
  The results of the comparison are shown in Figure \ref{fig:aRTgraphs},
  which follow the classical BBOB 2009 format, which includes the amount
  of effort devoted to finding a certain fitness level and time needed
  to do it. Figure \ref{fig:aRTgraphs} was generated by the postprocessing script from COCO \cite{hansen2016coco} used in the Black-box Optimization Benchmarking workshop series.   
  The EvoSpace and KafkEO results are shown side by side, only
  for the sake of comparison, since we are only interested for the time
  being in the baseline performance of the proof of concept.
